\chapter{Clouds and chemistry}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% author: Celine Mari, CNRS/LA (20/03/08)
% Adding chemistry in resolved clouds :  Maud Leriche (01/07/2011)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\minitoc
%
Clouds play an important role in atmospheric chemistry.
Convective clouds transport pollutants vertically from the boundary layer to 
the upper troposphere. Clouds and associated precipitations scavenge pollutants 
from the air. Once inside the cloud or rain water, some compounds 
dissociate into ions and/or react with the one another through aqueous 
chemistry.
Another important role for clouds is the removal of pollutants trapped in 
rain water and its deposition onto the ground. Clouds can also affect 
gas-phase chemistry by attenuating solar radiation below the cloud base,
which has a significant impact on the photolysis reactions. 
The model currently incorporates parameterizations for sub-grid convective 
clouds (precipitating and non-precipitating) and for grid-scale 
resolved clouds. 
%-------------------------------------
\section{Model description}
%-------------------------------------
The cloud scheme can be divided into two main components: the 
{\bf sub-grid} cloud model and the {\bf resolved} cloud model. 
For large horizontal grid resolutions, the grid size will be larger than the 
size of a typical convective cloud, requiring a parameterization for these
sub-grid clouds. The sub-grid cloud scheme simulates convective 
precipitating and non-precipitation clouds (see Chapter:Convection Scheme).
The second component of the cloud model considers clouds which occupy the 
entire grid cell and have been "resolved" by the model. The rate of change in 
pollutant concentrations ($s_i$) due to cloud processes is given by:
$$
\frac{\partial C_i}{\partial t} = \frac{\partial C_i}{\partial t}_{subcld} 
+ \frac{\partial C_i}{\partial t}_{rescld}
$$
%-------------------------------------
\section{Subgrid convective cloud scheme}
%-------------------------------------
%-------------------------------------
\subsection{Wet scavenging}
%-------------------------------------

Scavenging by subgrid wet convective updrafts is applied within the 
convective mass transport algorithm (see Chapter:Convection Scheme) in 
order to prevent soluble tracers from being transported to the top of the 
convective updraft and then dispersed on the grid scale. The transport model 
provide wet convective air mass fluxes through each grid level in the 
updraft. As air is lifted a distance $\Delta z$ from one level to the next, 
it loses a fraction $\rm F_i$ of soluble tracer $i$ to scavenging. This 
fraction depends on (1) the rate constant $k$ ($\rm s^{-1}$) for conversion 
of cloud condensate (including liquid and ice) to precipitation; 
(2) the fraction $f_{i,L}$ of tracer present in the liquid cloud condensate; 
(3) the fraction $f_{i,I}$ of tracer present in the ice cloud condensate; 
and (4) the retention efficiency $R_i$ of tracer in the liquid cloud condensate
as it is converted to precipitation ($R_i <$ 1 accounts for volatilization 
during riming). Thus the rate constant $k_i$ ($\rm s^{-1}$) for loss of 
tracer from the updraft is given by \citet{Mari2000}:
$$
k_i = (R_i f_{i,L} + f_{i,I}) k
$$
and the fraction $F_i$ of tracer scavenged as the air is lifted by $\Delta z$ 
is
$$
F_i = 1 - \exp\left[ -k_i \frac{\Delta z}{w} \right]
$$
where $w$ is the updraft velocity. The scavenged tracer is directly deposited 
to the surface, there can be no re-evaporation.
\subsubsection*{$\rm HNO_3$} 
$\rm HNO_3$ is 100\% in the cloud condensate phase ($\rm f_{i,L} + f_{i,I}=1$) 
and we assume $R_i = 1$, therefore $k_i = k$ \citep{Mari2000,Liu2001}.
\subsubsection*{Gases other than $\rm HNO_3$} 
For gases other than $\rm HNO_3$, a significant fraction of tracer may be in 
the gas phase so that $k_i < k$. The phase partitioning of the tracer depends 
on the cloud liquid water content, L ($\rm cm^3$water$\rm cm^{-3}$air), and the 
cloud ice water content W ($\rm cm^3$ice$\rm cm^{-3}$ air). 

Let $C_{i,G}, C_{i,L}, C_{i,I}, C_{i,T}$ represent the atmospheric mixing 
ratios of tracer in the gas, liquid, cloud condensate, ice cloud condensate, 
and all phases, respectively, so that 
$$
 C_{i,T} = C_{i,G} + C_{i,L} + C_{i,I}
$$
$$
f_{i,L} = \frac{C_{i,L}}{C_{i,T}} = \frac{\frac{C_{i,L}}{C_{i,G}}}{1 + \frac{C_{i,L}}{C_{i,G}} + \frac{C_{i,I}}{C_{i,G}}}
$$
$$
f_{i,I} = \frac{C_{i,I}}{C_{i,T}} = \frac{\frac{C_{i,I}}{C_{i,G}}}{1 + \frac{C_{i,L}}{C_{i,G}} + \frac{C_{i,I}}{C_{i,G}}}
$$
The ratio $C_{i,L}/C_{i,G}$ is obtained from Henry's law:
$$
\frac{C_{i,L}}{C_{i,G}} = K_i^* LRT
$$
where $K_i^*$ (M atm$^{-1}$) is the effective Henry's law constant including 
contributions from dissociated species in fast equilibrium with the dissolved 
tracer, $R = 8.32\times 10^{-2}$ atm M$^{-1}$ K$^{-1}$ is the ideal gas constant, and T is the local temperature in K. We calculate $K_i^*$ from the van't Hoff
equation: 
$$
K_i^* = K_{i 298}^* \exp\left[ -\frac{\Delta H^0_{i 298}}{R}\left(\frac{1}{T} - \frac{1}{T_0} \right) \right]
$$
where $T_0 = 298 K$. 

The retention efficiency $R_i$ is unity for all gases in a warm cloud
(T$\ge$268 K). Values of $R_i$ in a mixed cloud 
($\rm 248 < T < 268$ K) are 0.02 for $\rm CH_3OOH$ and $\rm CH_2O$ and 0.05 for $\rm H_2O_2$ \citep{Mari2000}.

The ratio $C_{i,I}/C_{i,G}$ for $\rm H_2O_2$ is obtained by assuming scavenging
by co-condensation: 
$$
\frac{C_{i,I}}{C_{i,G}}=
\frac{W}{C_{H_2O}}
\left( \frac{\alpha_{H_2O_2}}{\alpha_{H_2O}} \right) 
\left( \frac{M_{H_2O_2}}{M_{H_2O}} \right) ^{\frac{1}{2}}
$$
where $C_{H_2O}$ is the water vapor mixing ratio (to be calculated from 
saturation over ice at the local temperature), $\alpha_{H_2O_2}/\alpha_{H_2O}$=0.6 is the ratio of sticking coefficients on the ice surface, and $M_{H_2O_2}/M_{H_2O}$=1.9 is the ratio of molecular weights. 

For $\rm CH_3OOH$ and $\rm CH_2O$ scavenging by co-condensation is inefficient
and we assume $C_{i,I}/C_{i,G}$=0.

%-------------------------------------
\subsection{Lightning produced NOx}
%-------------------------------------
The mass flux formalism applied to the convective transport of a chemical 
compound $\overline{C}$, writes :
%
\begin{equation}\label{transport_convectif}
{\frac {\partial \overline C}{\partial t}}\bigg\vert_{convection} =
-\frac{1}{\rho A} \frac {\partial (MC)}{\partial z} 
- \overline{w} {\frac {\partial \overline C}{\partial z}}
\end{equation}
%
where $A$ is the grid mesh area, $\rho$ is the air density, $M$ is the mass flux
(in kg/s) and $\overline{w}$ is the environmental subsidence to compensate the
upward mass flux. $C$ is the mixing ratio of chemical compound in the 
convective cells. The mass flux term of Eq. (\ref{transport_convectif}) is
further decomposed into:
%
\begin{equation}\label{decomposition}
\frac {\partial (MC)}{\partial z} = \frac {\partial (M^u C^u)}{\partial z} + \frac {\partial (M^d C^d)}{\partial z}
\end{equation}
%
where the superscripts $u$ and $d$ refer to the "updrafts" and to the 
"downdrafts" components, respectively. The different mass flux divergences are 
expressed 
as:
%
\begin{equation}\label{div_mass_fluxes_u}
\frac{\partial}{\partial z} (M^u C^u) = \epsilon^u\overline{C}-\delta^u C^u
\end{equation}
\begin{equation}\label{div_mass_fluxes_d}
\frac{\partial}{\partial z} (M^d C^d) = \epsilon^d\overline{C}-\delta^d C^d
\end{equation}
%
where $\epsilon$ and $\delta$ are the parameterized entrainment and detrainment
rates, respectively. Selecting $C$ as the mixing ratio of nitrogen monoxide, 
$[NO]$, Eqs (\ref{div_mass_fluxes_u})-(\ref{div_mass_fluxes_d}) are modified to
include the internal LiNOx production rates \citep{Mari2006}:
%
\begin{equation}\label{div_flux_NO_u}
\frac{\partial}{\partial z} (M^u[NO]^u) = \epsilon^u\overline{[NO]}-\delta^u[NO]^u + (\overline{\rho} A) \frac{\partial [NO]^u}{\partial t}\bigg\vert_{LiNOx}
\end{equation}
\begin{equation}\label{div_flux_NO_d}
\frac{\partial}{\partial z} (M^d[NO]^d) = \epsilon^d\overline{[NO]}-\delta^d[NO]^d
\end{equation}
%\begin{equation}\label{div_flux_NO_d}
%{\partial\over\partial z} (M^d[NO]^d) = \epsilon^d\overline{[NO]}-\delta^d[NO]^d + (\overline{\rho} A) \frac{\partial [NO]^d}{\partial t}\bigg\vert_{LiNOx}
%\end{equation}
%
The two terms on the right hand side of 
Eqs (\ref{div_flux_NO_u}) and (\ref{div_flux_NO_d})
represent the subgrid scale
transport of NO.
Transport of NO 
is assumed to take place instantaneously during each model timestep. 
The third term is the LiNOx term to be parameterized.
It is worth noting that 
no a-priori vertical placement of LiNOx is necessary with this approach.
Once produced inside the convective column, NO molecules are redistributed 
by upward and downward transport and detrained in the environment. The 
vertical placement of LiNOx is a direct consequence of the redistribution by 
mass fluxes inside the convective scheme. 

The electrical activity in the thunderstorms is related to the vertical 
extension of the glaciated region where ice-ice particle rebounding collisions 
are efficient enough to explain the charging mechanisms 
\citep{Reynolds1957,Takahashi1978,Saunders1992}. 
A growing electrical 
field then results from the organization of dipolar, tripolar or even 
more complex charge 
structures at storm scale \citep{Rust2002,Rust1996,Stolzenburg2002,Barthe2005}. 
The electrical field is broken down by a partial 
neutralization of the electrical charges. This is realized by a repetitive 
triggering of intra-cloud (IC) and cloud-to-ground (CG) flashes. The flashes 
lead to the formation of NO in the lightning channels after dissociation
of air molecules at high temperature followed by a rapid cooling. 
According to 
the statistical regression formula of \citet{Price1992}, 
the total lightning
frequency over land and ocean, $f_f$ can be grossly estimated from 
mean cloud morphological parameters: 
%
\begin{equation} \label{eqn : frequence}
f_f=3.44 \times 10^{-5}{H_{ct}}^{4.9} \quad
\end{equation}
over land and 
\begin{equation} \label{eqn : frequence2}
f_f=6.40 \times 10^{-4}{H_{ct}}^{1.73} \quad
\end{equation}
over ocean.
$H_{ct}$ is the cloud top height of the convective cells (in km).

\citet{Price1993} proposed the following polynomial relationship
between the thickness of the cold icy cloud ($H_{fr}$ in km) and the 
IC/CG ratio, $\beta$:
\begin{equation} \label{eqn : rapport}
\beta =0.021H_{fr}^4-0.648H_{fr}^3+7.493H_{fr}^2-36.54H_{fr}+63.09 \quad
\end{equation}
with $\rm 1<\beta<50$.
%
A scaling factor 
$c_{pr}=0.97241 \; \exp(0.048203\times \Delta lat \Delta lon)$ is introduced 
by \citet{Price1994} to adapt $f_f$ to different mesh sizes in 
interval of latitude ($\Delta lat$) and longitude ($\Delta lon$) given in 
degree. 

The combination of Eqs (\ref{eqn : frequence})-(\ref{eqn : rapport}), leads to
the final expression of the LiNOx production rates to be inserted in
Eq (\ref{div_flux_NO_u}). It can be written in condensed form :
\begin{equation}\label{LiNOx_rate_ud}
{\frac {\partial (NO)^{u}}{\partial t}} \bigg\vert_{LiNOx} = \frac{\beta f_f}{1+\beta}\times P(IC) + \frac{f_f}{1+\beta}\times P(CG)
\end{equation}
where the value of the mean production rate per CG and IC namely, 
$P(CG) = 6.7\times 10^{26}$ of NO molecules, and 
$P(IC) = 6.7\times 10^{25}$ of NO molecules, of
\citet{Price1997} have been retained. 
It is worth noting that recent studies based on airborne observations and 
cloud scale modeling found that intracloud flashes are likely to be as 
effective 
in producing NO as cloud-to-ground flashes 
\citep{DeCaria2000,Fehr2004,Ridley2005}. 
It is also important to note that the estimates for the lightning 
NOx production based on \citet{Price1992,Price1994,Price1997}
are on the high end of current estimates in the  literature
\citep{Labrador2005}. 

The practical implementation of the LiNOx parameterization is based on critical 
vertical levels defined in the deep convection scheme. IC flashes are equally 
distributed between the cloud top and the freezing levels. The CGs are located 
between the -10$^{\circ}$C level (or the level of free sink if below) and the 
ground. The NO production in flashes 
is assumed to be proportional to air density following 
\citet{Goldenbaum1993}. 
Recent studies have shown that unimodal or bimodal distributions would be more 
realistic than uniform distributions as discussed in 
\citet{MacGorman1998,DeCaria2000,DeCaria2005}. 

%-------------------------------------
\section{Resolved Cloud Scheme}
%-------------------------------------
%-------------------------------------
 \subsection{Cloud chemistry module}
%-------------------------------------
The cloud chemistry module allows solving both gaseous and aqueous phases chemistry 
as well as mass transfer reactions between gas and liquid phases and microphysical 
transfers of chemical species \citep{Leriche2013}. In nested mode, to avoid time computing, 
the cloud chemistry module can be available in the innest domain, the one at high resolution, 
while the coupling models (father models) treat the gas phase chemistry only. This possibility 
is obviously available in one--way nesting mode. The concentrations of chemical species $i$ 
in gas phase and in aqueous phase (cloud water and rainwater) are solution of 
the set of differential equations:

\begin{equation}
\frac{\partial C_{g}^{i}}{\partial t}_{rescld}=\frac{\partial C_{g}^{i}}{\partial t}_{chem}+\frac{\partial C_{g}^{i}}{\partial t}_{emis}+\frac{\partial C_{g}^{i}}{\partial t}_{dep\_sec}+\frac{\partial C_{g}^{i}}{\partial t}_{mic}
\end{equation}
\begin{equation}
\frac{\partial C_{w}^{i}}{\partial t}_{rescld}=\frac{\partial C_{w}^{i}}{\partial t}_{chem}+\frac{\partial C_{w}^{i}}{\partial t}_{mic}
\end{equation}

The equation for gas phase is the same than without cloud except for the last term, which represents the release of soluble gases from the aqueous phase when evaporation occurs or when freezing or riming occurs. 
In equation for liquid phase, $w$ stands for either cloud water or rainwater. The term $mic$ represents the sources and sinks due to cloud microphysical processes. These terms depend on the cloud microphysics scheme used.

%-------------------------------------
 \subsection{Chemical kinetic scheme}
%-------------------------------------
The evolution of the chemical concentrations (in mol per volume of air ) in the gas phase and in liquid phases denoted by the term $chem$ above of a chemical species $i$ are given by the generic set of differential equations:

\begin{equation}
  \frac{\partial C_{g}^{i}}{\partial t}_{chem}=P_{g}^{i}-D_{g}^{i}C_{g}^{i}-k_{tw}^{i}\left (   L_{w}C_{g}^{i}-\frac{C_{w}^{i}}{H_{eff}^{i}RT} \right )
  \label{toto1}
\end{equation}
\begin{equation}
  \frac{\partial C_{w}^{i}}{\partial t}_{chem}=P_{w}^{i}-D_{w}^{i}C_{w}^{i}+k_{tw}^{i}\left ( L_{w}C_{g}^{i}-\frac{C_{w}^{i}}{H_{eff}^{i}RT} \right )
  \label{toto2}
\end{equation}

Subcripts $g$ and $w$ are, respectively, for the gas phase and for liquid phases with $w$ stands either for the cloud droplets or for the raindrops. Terms $P$ and $D$ are the production terms in mol per volume of air s$^{-1}$ and the destruction terms in s$^{-1}$. $L$ is the liquid water volume ratio (volume of the drops by volume of air) computed from the liquid water mixing ratio using air and water density. $H_{eff}$ is the effective Henry's law constant in mol atm$^{-1}$. $T$ is the temperature in K and $R$ = 0.08206 atm mol$^{-1}$ K$^{-1}$ is the universal gas constant. The rate constant of transfer between the gas phase and the aqueous phase $k_{tw}$ in s$^{-1}$ is the inverse of the characteristic times for gaseous diffusion and for the interfacial mass transport according to \citet{Schwartz1986}:

\begin{equation}
k_{tw}^{i}=\left ( \frac{a_{w}^{2}}{3D_{g}^{i}}+\frac{4a_{w}}{3\overline{v}^{i}\alpha^{i}} \right )^{-1}
\end{equation}

the parameter $a_{w}$ is the cloud droplet or raindrop radius in m, $D_{gas}$ is the gaseous diffusion coefficient , $\overline{v}$ is the mean quadratic speed in m s$^{-1}$ of the soluble species and $\alpha$ is the accommodation coefficient of the soluble species. The value of the drop radius for cloud droplets and raindrops comes from the microphysical scheme. 
In equations (\ref{toto1} and \ref{toto2}), the two first terms at the right hand side are the chemical production and destruction terms and the last term is the reversible uptake of soluble gases inside cloud droplets or raindrops. The complete set of reactions of the cloud chemistry module is described in \citet{Leriche2013}. This chemical mechanism called ReLACS-AQ has been developped based upon the gas phase mechanism ReLACS \citep{Crassier2000} and includes 25 additional chemical species in liquid phases (cloud water and rainwater) for 30 photo-chemical reactions and 20 reactions of exchange between gas and aqueous phase. Thus, 91 prognostics chemical species are considered in the cloud chemistry scheme. The aqueous phase reactivity describes the reactivity of $HO_{x}$, of formadehyde and formic acid and of sulfur and nitrogen leading to the formation of sulfuric and nitric acid. All kinectic and thermodynamic data used are given in \citet{Leriche2013}.

%-------------------------------------
\subsection{Microphysical transfer terms}
%-------------------------------------
Once dissolved or produced in the drops, the aqueous chemical species is redistributed between the cloud droplets and the raindrops by the cloud microphysical processes. Three processes are implied in the microphysical transfer of aqueous chemical species in cloud: autoconversion, accretion and sedimentation, which lead ultimately to the wet deposition of the soluble species by rain. These microphysical transfer terms are computed following:
\begin{equation}
\frac{\partial C_{c}^{i}}{\partial t}_{AUTO}=-\frac{\partial C_{r}^{i}}{\partial t}_{AUTO}=-C_{c}^{i}\frac{1}{r_{c}}\frac{\partial r_{c}}{\partial t}_{AUTO}
\end{equation}
\begin{equation}
\frac{\partial C_{c}^{i}}{\partial t}_{ACCR}=-\frac{\partial C_{r}^{i}}{\partial t}_{ACCR}=-C_{c}^{i}\frac{1}{r_{c}}\frac{\partial r_{c}}{\partial t}_{ACCR}
\end{equation}
\begin{equation}
\frac{\partial C_{r}^{i}}{\partial t}_{SEDI}=-C_{r}^{i}\frac{1}{r_{r}}\frac{\partial r_{r}}{\partial t}_{SEDI}=-C_{r}^{i}\frac{1}{r_{r}}\frac{\partial }{\partial z}F_{SEDI}\left ( r_{r} \right )
\end{equation}
$r_{c}$ and $r_{r}$ are the mass mixing ratios of the cloud droplets and of the raindrops, respectively. Subscripts AUTO, ACCR and SEDI are respectively for autoconversion process, accretion process and the sedimentation of the raindrops. $F_{SEDI}$ is the sedimentation flux of the rain mixing ratio. The sedimentation flux and the tendencies of cloud droplets mixing ratio for autoconversion and accretion processes depend on the cloud microphysical scheme. 

%-------------------------------------
\subsection{pH diagnostic equation}
%-------------------------------------
As the solubility of some important gaseous pollutants depends on the pH of the drops, and the chemical reactivity in the aqueous phase as well, it is crucial to predict the evolution of pH in cloud droplets and in raindrops. The pH is the solution of the electroneutrality equation which expresses the exact balance between the cations and the anions in an aqueous solution. By using a simplified form of the electroneutrality equation, the pH equation leads to a high order polynomial equation of the concentration of the protons $\rm{pH} = log_{10}(H^{+})$. The physical root is found using the Laguerre method. In contrast with other technique \citep{Barth2007}, the pH is computed exactly and updated before each substep of the Rosenbrock algorithm to ensure numerical accuracy and stability of the aqueous phase reactions. Most of the coefficients of the pH polynomial equation are complex expressions containing reaction transfer rates. These expressions are automatically simplified using a formal calculus algorithm and not by hands to avoid errors.


%-------------------------------------
\subsection{Extension to the ice phase}
%-------------------------------------
In mixed phase clouds, additional processes for chemical species concentrations evolution appear including direct gas uptake by ice crystals, partitioning during freezing or riming of liquid hydrometeors and surface and bulk reactions in/on ice hydrometeors. Several studies have shown \citep{Yin2002,Long2010} that the main process to be considered in the evolution of chemical species concentrations in mixed phase clouds is the retention of soluble gases when liquid hydrometeors freeze/rime. In particular, while the direct uptake inside growning ice crystals by vapor deposition is almost always negligeble, this process is not implemented. The freezing and the riming of liquid drops transfer aqueous phase chemical species from liquid hydrometeors (cloud dropplets and raindrops) to precipitating iced hydrometeors i.e. snow, graupel and hail in Meso--NH. In order to limit the number of prognostic variables, the concentrations of chemical species in snow, graupel and hail have been treated globally as:
\begin{equation}
C_{ice}^{i}=C_{snow}^{i}+C_{grau}^{i}+C_{hail}^{i}
\end{equation}
The additional microphysical transfer terms due to the retention of soluble gases during freezing/riming are computed following:

\begin{equation}
\frac{\partial C_{g}^{i}}{\partial t}_{FZ}=\left ( 1-RET \right )C_{w}^{i}\frac{1}{r_{w}}\frac{\partial r_{w}}{\partial t}_{FZ}
\end{equation}
\begin{equation}
\frac{\partial C_{w}^{i}}{\partial t}_{FZ}=-C_{w}^{i}\frac{1}{r_{w}}\frac{\partial r_{w}}{\partial t}_{FZ}
\end{equation}
\begin{equation}
\frac{\partial C_{ice}^{i}}{\partial t}_{FZ}=RETC_{w}^{i}\frac{1}{r_{w}}\frac{\partial r_{ice}}{\partial t}_{FZ}=RETC_{w}^{i}\frac{1}{r_{w}}\left ( \frac{\partial r_{snow}}{\partial t}_{FZ}+\frac{\partial r_{grau}}{\partial t}_{FZ}+\frac{\partial r_{hail}}{\partial t}_{FZ} \right )
\end{equation}
\begin{equation}
\frac{\partial C_{ice}^{i}}{\partial t}_{SEDI}=-C_{ice}^{i}\frac{1}{r_{ice}}\frac{\partial r_{ice}}{\partial t}_{SEDI}=-C_{ice}^{i}\left ( \frac{1}{r_{snow}}\frac{\partial r_{snow}}{\partial t}_{SEDI}+\frac{1}{r_{grau}}\frac{\partial r_{grau}}{\partial t}_{SEDI}+\frac{1}{r_{hail}}\frac{\partial r_{hail}}{\partial t}_{SEDI} \right )
\end{equation}

$FZ$ refers to freezing and riming processes. $RET$ is the retention coefficient. A zero value for $RET$ means that the soluble gas is completely released to the gas phase and is not retained in the ice phase at all. A value of 1 means that the soluble gas is completely retains in the ice phase. The last term $SEDI$ refers to the sedimentation of graupel, snow and hail, which contributes to wet deposition.

%-------------------------------------
 \subsection{Lightning produced NOx at cloud scale}
%-------------------------------------
To be implemented
%\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%% BIBLIOGRAPHY %%%%%%%%%%%%%%%%%%%%%%%%
\begin{btSect}{4-3-CloudGas}
\section{References}
\btPrintCited
\end{btSect}
%%%%%%%%%%%%%%%%%%%%%%%%%%%% BIBLIOGRAPHY %%%%%%%%%%%%%%%%%%%%%%%%

